---
title: "Fit Trends for 2023 (Transform data - types as columns)"
author: "Julia Wu"
date: "2023-11-01"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Used libraries

```{r load}

library(tidyverse)
library(kableExtra)
library(reshape2)

```

## Import cleaned data

```{r import}

df <- read.csv("fitness trends 2023 (cleaned).csv")

df %>% 
  kable() %>%
  kable_styling(bootstrap_options= c("striped", "hover")) %>%
  scroll_box(width="100%", height= "500px")

```

## Transform data

### Initiate a dataframe with default value 0 for all types of training

### Types as Columns

```{r new_frame}

types_all <- as.list(df[, 2:ncol(df)])
types_uni <- sort(unique(unlist(types_all, recursive=FALSE)))

Types <- c(types_uni)
new_trends <- data.frame(c(2007:2023))

new_trends[c(Types)] <- 0

colnames(new_trends)[1]<- "Year"

new_trends %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover")) %>% 
  scroll_box(width="100%", height="300px")

```

### Convert ranking to popularity in a time series

```{r convert}

for(Type in Types){
  #print(Type)
  
  for(year in 2007:2023){
    for(rank in colnames(df)[2:ncol(df)]){
      matches <- df[(df[rank]==Type) & (df['Year']==year),]
      
      if(nrow(matches)>0){
        #print(matches)
        
        index <- which(matches==Type, arr.ind=FALSE)
        #print(index)
        
        popularity <- 12-index
        #print(popularity)
        
        new_trends[new_trends['Year']==year, Type] <- popularity
      }
    }
  }
}
  
new_trends %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover", "responsive")) %>% 
  scroll_box(height="300px")

```

### Save as a .csv

```{r save}

write_csv(new_trends, "fitness trends 2023 (transformed_type).csv")

```

### Read from a .csv

```{r read}

trends_con <- read.csv("fitness trends 2023 (transformed_type).csv", check.names=FALSE)

trends_con %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover", "responsive")) %>% 
  scroll_box(height="300px")

```

## Filter interesting data points

### Find popularity of all types of training for recent 5 years

```{r recent}

trends_con_recent <- trends_con[(trends_con['Year']>=2019), ]

trends_con_recent %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover")) %>% 
  scroll_box(width="100%")

```

### Types of training with a total popularity for recent 5 years \> 15, and in 2023 \> 4

```{r top}
trends_con_recent_top <-trends_con_recent[
  colSums(trends_con_recent, na.rm=TRUE)>15 & 
  trends_con_recent[trends_con_recent['Year']==2023]>4]

trends_con_recent_top %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover", "responsive"))

```

## Melt data and convert data type as needed for time series graph

```{r final}

data_final <- melt(trends_con_recent_top, id.vars='Year')

data_final %>% 
  kable() %>% 
  kable_styling(bootstrap_options=c("striped", "hover")) %>% 
  scroll_box(width="100%", height="300px")

```

## Draw plots

```{r plot}

ggplot(data_final, aes(x=Year, y=value, col=variable))+
  geom_line()+
  scale_x_continuous(breaks=pretty(data_final$Year, n=4))+
  scale_y_continuous(breaks=pretty(data_final$value, n=5))+
  
  labs(x="Year", y="Popularity", col="Types")+
  ggtitle("Types as Columns")+
  theme(plot.title=element_text(hjust=0.5))

```

#### \-\--

#### References:

[@thompson2023]
